---
title: "Homework Assignment 4"
author: "Parker Reedy"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---


```{r setup, echo=FALSE}
library(knitr)
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)

## indents are for indenting r code as formatted text
## They may need to be adjusted depending on your OS
# if your output looks odd, increase or decrease indent
indent1 = '    '
indent2 = '        '
indent3 = '            '
```

```{r, echo=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(dendextend)
library(dplyr)
```

1. 
```{r, message=FALSE}
leukemia_data <- read_csv("leukemia_data.csv")
```
a. From the table, we can see that the least common subtype in the data is "BCR-ABL."
```{r}
leukemia_data <- leukemia_data %>% mutate(Type = as.factor(Type))
leukemia_table <- table(leukemia_data$Type)
leukemia_table
```

b. you would need 201 Principal Components to explain 90% of the total variation in the data. 
```{r}
pca_leukemia <- prcomp(select(leukemia_data, -Type), scale=TRUE, center=TRUE)

pr.var = pca_leukemia$sdev^2

pve = pr.var/sum(pr.var)

plot(pve, xlab='Principal Component', ylab='Proportion of Variance Explained', ylim=c(0,1),type='b')

plot(cumsum(pve), xlab='Principal Component', ylab='Cumulative Proportion of Variance Explained', ylim=c(0,1),type='b')

which(cumsum(pve) > .90)[1]

```


c. The T-ALL group is most clearly separated from the others along the PC2 axis. The 6 genes with the largest absolute weights are SEMA3F, CCT2, LDHB, COX6C, SNRPD2, and ELK3.
```{r}
rainbow_colors <- rainbow(7)
plot_colors <- rainbow_colors[as.factor(leukemia_data$Type)]

plot(pca_leukemia$x[, 1], pca_leukemia$x[, 2], col = plot_colors, cex = 0, main = 'PCA of Leukemia Data', xlab='PCA 1', ylab = 'PCA 2')
text(pca_leukemia$x[, 1], pca_leukemia$x[, 2], labels = leukemia_data$Type, col = plot_colors, cex = .7)

head(names(sort(abs(pca_leukemia$rotation[, 1]), decreasing=TRUE)))
```

d. The third PC does seem to be better at discriminating between leukemia types because in this plot, the leukemia types are more grouped together between their individual types.
```{r}
plot(pca_leukemia$x[, 1], pca_leukemia$x[, 3], col = plot_colors, cex = 0, main = 'PCA of Leukemia Data', xlab='PCA 1', ylab = 'PCA 3')
text(pca_leukemia$x[, 1], pca_leukemia$x[, 3], labels = leukemia_data$Type, col = plot_colors, cex = .7)
```

e.
```{r, fig.height = 15}

leukemia_subset <- filter(leukemia_data, Type %in% c('T-ALL','Hyperdip50','TEL-AML1'))

distance_matrix <- dist(select(leukemia_subset, -Type), method = 'euclidean')
leukemia.hclust <-  hclust(distance_matrix)

dend1 <- as.dendrogram(leukemia.hclust)
dend1 = color_branches(dend1, k=3)
dend1 = color_labels(dend1, k=3)
dend1 = set(dend1, 'labels_cex', 0.3)
dend1 = set_labels(dend1, labels=leukemia_subset$Type[order.dendrogram(dend1)])

plot(dend1, horiz=T, main = "Dendrogram colored by 3 clusters", cex = 0.5)



dend2 <- as.dendrogram(leukemia.hclust)
dend2 = color_branches(dend2, k=5)
dend2 = color_labels(dend2, k=5)
dend2 = set(dend2, 'labels_cex', 0.3)
dend2 = set_labels(dend2, labels=leukemia_subset$Type[order.dendrogram(dend2)])

plot(dend2, horiz=T, main = "Dendrogram colored by 5 clusters", cex = 0.5)

```













